Mariculture farms in the United States

### 2021-12-14 marine/united_states/US_mapping.Rmd
updated us mapping exploration script

---
title: "Location of United States Aquaculture Farms"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "10/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script maps all known US mariculture farms

@@

```{r, include = FALSE}
library(raster)
library(tidyverse)
library(stringr)
library(PBSmapping)
library(maptools)
library(rgdal)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(tmap)
library(leaflet)
library(ggplot2)
library(maps)
library(mapdata)

@@

# Raster

Standard raster used for this analysis

@@

## US Northeast

@@

## Maine:

  ggplot() +
  geom_sf(data = maine_aq_lease_active, col = "red") + 
  geom_polygon(data=maine, aes(x=long, y=lat, group=group), fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Maine data", caption = "Comprehensive, up-to-date (2021)") +
    geom_sf(data = maine_aq_lp_active, col = "blue", alpha = 0.1, size = 0.5)
```

## Alaska:

## California:

## New Hampshire:

## Florida:

@@

## Alabama:

## Mississippi:

 - Can't find much information for Mississippi.

## Louisana:

@@

## Texas:

@@

## Oregon:

## Connecticut:

However, worst case, we could use the data from the Northeast ocean council dataset:

```{r}

Another map, this is probably what we need (seaweed and shellfish): https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

## Washington:

@@

## All states:

## North Carolina:

 - https://deq.nc.gov/about/divisions/marine-fisheries/licenses-permits-and-leases/shellfish-lease-and-franchise#stats-and-figures
 - Comprehensive for shellfish... just need to figure out how download it. 
 - I believe Shellfish is pretty much the only mariculture that occurs in NC. 
 - Has number totals of leases 
 - Will likely need to send an email.

## Virginia:

 - https://mrc.virginia.gov/links.shtm
 - They have oyster lease and application data, which *I think* is the only type of aquaculture there. The map isn't downloadable from what I can tell though, so I bet we will have to email to access.
 - They also have a phone app.

## Maryland:

 - They have individual lease application maps in PDFs, but they would take forever to go through. They must have some sort of internal dataset where they store the approved lease info?: https://dnr.maryland.gov/fisheries/pages/aquaculture/applications.aspx

## Delaware:

@@

## South Carolina:

## Georgia:

@@

## Massachusetts:

https://www.mass.gov/service-details/aquaculture
  - "Massachusetts has 378 shellfish leases. Shellfish farms are a source of healthy protein and can contribute to healthy marine ecosystems by cleaning the water column, providing habitat for other sea life, and helping protect shorelines from storm surges and sea level rise."

## Rhode Island:

 - Lease map here: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f
 - No download option.

## Hawaii

## New Jersey

 - this has maps of all shellfish leases in 2020.. but they are all pdf.

## New York

Suffolk County

 - Lease mapper: https://gis.suffolkcountyny.gov/portal/apps/View/index.html?appid=6ae708a346f340ada850544fe25ce0fb
 - Its not totally unreasonable to think that this is the only place in New York that does mariculture? Since its coast is so small.

 - Shellfish map for NY and CT, this seems like what we would need: https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

>>>>

---
title: 'Mariculture farms in the United States'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

Exploring maps and data for all known US mariculture farms

@@

```{r, include = FALSE}
library(raster)
library(tidyverse)
library(stringr)
library(PBSmapping)
library(maptools)
library(rgdal)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(tmap)
library(leaflet)
library(ggplot2)
library(maps)
library(mapdata)
library(readr)

@@

state <- map_data("state")
```

@@

## Overview

There are 23 states with marine coasts.

**There is comprehensive downloadable data for 5 states: **

 - Maine, Alaska, New Hampshire, New York(?), Connecticut(?)

**There are publicly available map viewers (leases or farm locations) which seem to be comprehensive, but are NOT downloadable for 4 states:**

 - Alabama, North Carolina, Virginia, Rhode Island

**There is general info on lease holders, what they farm, and what municipality or city they are in for 6 states:**

 - California: general areas with acreage for commercial farms
 - Oregon: map of places were commercial harvest/growing is allowed
 - Maryland: Have INDIVIDUAL pdf maps of all lease applications and if they were accepted or not...
 - New Jersey: Have pdf maps of all existing shellfish leases and those issued in 2020
 - Massachusetts: Have number of growers per municipality with acreage and species type; also a subset of farms are available in the NE data viewer.
 - South Carolina: Apparently have 9 oyster farms as of 2021

**Subset of farms downloadable for 1 state:**

 - Florida: A map of mariculture farms which are at risk of oil spills

**Couldn't really find any information on 5 states:**

 - Hawaii, Washington, Texas, Louisiana, Mississippi

**Probably no mariculture for 2 states:**

 - Delaware: Commercial harvest is illegal
 - Georgia: From what I could tell, oyster farming was illegal here as recent as 2020

## Comprehensive, downloadable

### Maine

ggplot() +
  geom_sf(data = maine_aq_lease_active, col = "red") + 
  geom_polygon(data=maine, aes(x=long, y=lat, group=group), fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Maine data", caption = "Comprehensive, up-to-date (2021)") +
    geom_sf(data = maine_aq_lp_active, col = "blue", alpha = 0.1, size = 0.5)

### Alaska

### New Hampshire

### Connecticut

From the Northeast ocean council dataset:

Another map combined with NY, this is what we need (seaweed and shellfish): https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

### New York

 - Lease mapper for Suffolk County: https://gis.suffolkcountyny.gov/portal/apps/View/index.html?appid=6ae708a346f340ada850544fe25ce0fb

 - Shellfish map for NY and CT, this seems like what we would need: https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46
 - Pretty easy to pull lat/lon coordinates from the attribute table on the web map.
 - CT has species and culture type. NY only has shellfish on the map

## Maps with no download

### Alabama

### North Carolina

 - https://deq.nc.gov/about/divisions/marine-fisheries/licenses-permits-and-leases/shellfish-lease-and-franchise#stats-and-figures
 - Comprehensive for shellfish... just need to figure out how download it. 
 - I believe Shellfish is pretty much the only mariculture that occurs in NC. 
 - Has number totals of leases 
 - Will likely need to send an email to NCDEQ

### Virginia

 - https://mrc.virginia.gov/links.shtm
 - They have oyster lease and application data, which *I think* is the only type of aquaculture there. The map isn't downloadable from what I can tell though, so I bet we will have to email to access.
 - They also have a phone app.
 - web.info@mrc.virginia.gov

### Rhode Island

 - Lease map here: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f
 - No download option.
 - Julia Livermore: julia.livermore@dem.ri.gov

## General lease info, areas, etc

### California

### Oregon

### Maryland

 - They have individual lease application maps in PDFs, but they would take forever to go through. They must have some sort of internal dataset where they store the approved lease info?: https://dnr.maryland.gov/fisheries/pages/aquaculture/applications.aspx
 - Maude Morris: maudee.morris@maryland.gov

### South Carolina

### New Jersey

 - this has maps of all shellfish leases in 2020.. but they are all pdf.
 - Amanda Wenczel: amanda.wenczel@ag.nj.gov

### Massachusetts

https://www.mass.gov/service-details/aquaculture
  - "Massachusetts has 378 shellfish leases. Shellfish farms are a source of healthy protein and can contribute to healthy marine ecosystems by cleaning the water column, providing habitat for other sea life, and helping protect shorelines from storm surges and sea level rise.
  - Sean Bowen: Sean.Bowen@mass.gov

## Subset available

### US Northeast dataset

@@

### Florida

@@

### All states

## Not much info

### Mississippi

 - Can't much info on Mississippi mariculture..

### Louisana

@@

### Texas

@@

### Washington

@@

### Hawaii

## No mariculture?

### Delaware

@@

### Georgia

@@

## Species/production breakdown for US

Take a look at the species in the US that FAO reports

 - These totals really don't match what NOAA reports in their 2019 report... They report 44094 metric tonnes of marine aquaculture production for 2018... versus 203687 from FAO. I think this has to do with NOAA reporting meat weight and FAO reporting live weight (includes shells)
 - https://media.fisheries.noaa.gov/2021-05/FUS2019-FINAL-webready-2.3.pdf?null=
 - NOAA also reports that the Gulf region produces the most marine aquaculture by volume.. which seems problematic for us considering we have very poor data for the Gulf states.
```{r, message=FALSE}
fao_data_all <- read_csv("../../data/fao_aquaculture_clean.csv")

fao_data_us <- fao_data_all %>%
  filter(iso3c == "USA")

fao_data_us_group <- fao_data_us %>%
  filter(year == 2018)

marine <- fao_data_us_group %>%
  filter(environment == "Marine") %>%
  filter(!is.na(value))

# unique(marine$species)

# 16 marine spp
#  [1] "Atlantic salmon"             "Blue mussel"                 "Northern quahog(=Hard clam)"
#  [4] "Sand gaper"                  "American cupped oyster"      "Red drum"                   
#  [7] "Whiteleg shrimp"             "Longfin yellowtail"          "Abalones nei"               
# [10] "Butter clam"                 "Cupped oysters nei"          "Japanese carpet shell"      
# [13] "Marine fishes nei"           "Pacific cupped oyster"       "Pacific geoduck"            
# [16] "Pacific littleneck clam"

## there were 4 species with NA production (so actually 20 marine species total)
marine_na <- fao_data_us_group %>%
  filter(environment == "Marine") %>%
  filter(is.na(value))
#unique(marine_na$species)
# [1] "Cockles nei"          "European flat oyster" "Olympia oyster"       "Pacific horse clam"

# sum(marine$value, na.rm = TRUE)

# 203687 total tonnes for marine.. does this seem correct? 
```
